---
title: "ESM 201 HW 1"
author: "Sara Orofino"
date: "1/30/2019"
output:
  pdf_document: default
  word_document: default
  always_allow_html: yes
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(kableExtra)
library(knitr)
library(stargazer)
library(tinytex)
library(captioner)
library(latex2exp)

population <- read_csv("assig1_data.csv")

pop <- population %>% 
  mutate(pop_mil = population/1000000)
```

```{r fig_caps, include=FALSE}
# set up for figure captions:
fig_nums <- captioner()

fig_nums("mex_percap", "Mexico's Actual vs. Calculated Per Capita Rate of Change. The graph depicts the actual vs. calculated per capita rate of change for Mexico's population between 1963 and 2003. The trendline shows the linear relationship between actual and calculate values for the per capita rate of population change.")

fig_nums("all_percap", "Rate of Population Change by Country. The per capita rate of change of the population is shown against the population size in millions of people for each of the five countries.")

fig_nums("all_color", "Rate of Population Growth by Country Over Time. The per capita rate of change of the population is shown against the population size in millions of people for each of the five countries. The color of the lines represents the progression through time and size of the points depicts the food calories per person per day.")

```


###Question 1 Calculate $\frac{dN}{Ndt}$ 

####a. Describe how annual population  are used to calculate $\frac{dN}{Ndt}$   

$\frac{dN}{Ndt}$ is the per capita rate of change of the population. Annual population data can be used to calculate the per capita rate of change (r) by taking the natural log of the number of individuals in the population at time t and divide by the number of individuals in the population at time t+1, multiplying it by $\frac{1}{(t+1)-t}$. In this case, the per capita rate of change would be the natural log of the number of individuals in the population in 1964 divided by the number of individuals in the population in 1963. For the difference from 1964 to 1963, the final term $\frac{1}{(t+1)-t}$ would be 1 and is not necessary to include in the equation. 

####b. Equations for calculating $\frac{dN}{Ndt}$  

* Brazil - $r = ln\left(\frac{81972001}{79602001}\right)$  

* India -  $r = ln\left(\frac{486639001}{476632001}\right)$   

* Japan -  $r = ln\left(\frac{96959001}{95929001}\right)$ 

* Mexico - $r = ln\left(\frac{43052001}{41715001}\right)$  

* South Korea - $r = ln\left(\frac{27767001}{27138001}\right)$    

```{r percap_ratechange, include=FALSE}

rates <- pop %>% 
  filter(year %in% c(1963,1964)) %>% 
  select(country, year, pop_mil) %>% 
  spread(year, pop_mil) %>% 
  magrittr::set_colnames(value = c("country", "n0", "nt")) %>% 
  mutate(dNNdt_calc = log(nt/n0),
         dNNdt = c(0.029338507, 0.020777870, 0.010679874, 0.031547910, 0.022913305))


```

  
    
```{r table1, echo=FALSE}

rate_table  <- rates %>%
  select(country, dNNdt_calc, dNNdt)

kable(rate_table, digits = 4, col.names = c("Country",  "Calculated dNNdt", "Actual dNNdt"),
      caption = "Calculated vs. Actual Per Capita Rate of Population Change by Country") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "bordered")) 

```
  
  
   
####c. Calcuate $\frac{dN}{Ndt}$ for Mexico for all years 1963-2004

```{r mexico, echo=FALSE, message=FALSE, warning=FALSE}

mexico <- read_csv("mexico_data.csv") %>% 
  mutate(pop_mil = population/1000000) %>% 
  select(country, year, pop_mil, dNNdt, dNNdt_calc) %>% 
  ggplot(aes(x = dNNdt_calc, y = dNNdt)) +
  geom_point() +
  geom_smooth(method="lm", se=FALSE, size = 0.5, color = "midnightblue") +
  labs(x = "Calculated Per Capita Rate of Change", 
       y = "Actual Per Capita Rate of Change") +
  scale_x_continuous(limits = c(0.010, 0.0325), breaks = seq(0.010,0.030, by=0.005),
                     expand = c(0,0)) +
  scale_y_continuous(limits = c(0.010,0.0325),breaks =  seq(0.010,0.030, by=0.005),
                     expand  = c(0,0)) +
  theme_light()

mexico
```
  
`r fig_nums("mex_percap")`    

\pagebreak


###Question 2 Graph $\frac{dN}{Ndt}$ vs. N~t~

####a. For Each Country Plot $\frac{dN}{Ndt}$ vs. N~t~

```{r pop_graphs, echo=FALSE}

ggplot(pop, aes(x = pop_mil, y = dNNdt)) +
  geom_line() +
  facet_wrap(~country, scales = "free") +
  labs(x = "Population Size (Millions of People)",
       y = "Per Capita Rate of Change") + 
  theme_light() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  
  
```
  
`r fig_nums("all_percap")`   


####b. Estimating Carrying Capacity 

```{r k, include=FALSE, message=FALSE, warning=FALSE}
# estimate carrying capacity for South Korea and Brazil 

# South Korea:

skorea_carry <- population %>% 
  filter(country == "South Korea") %>% 
  lm(population ~ dNNdt, data = .)

summary(skorea_carry)
# Equation: Population = 5.282x10^7 - 1.054x10^9(dNNdt)
# dNNdt = 0 at k --> K = 5.282x10^7

# Brazil:
brazil_carry <- population %>% 
  filter(country == "Brazil") %>% 
  lm(population ~ dNNdt, data = .)

summary(brazil_carry)
# Equation: Population = 2.607x10^8 - 6.392x10^9(dNNdt)
# dNNdt = 0 at k --> K = 2.607x10^8

```

  
**Carrying Capacities:**
  
- **Brazil:**   
$Population = 2.607*10^8 - 6.392*10^9(\frac{dN}{Ndt})$  
Carrying capacity (K) = population when $\frac{dN}{Ndt} = 0$   
$K = 2.607*10^8 individuals$
  
- **South Korea:**   
$Population = 5.282*10^7 - 1.054*10^9(\frac{dN}{Ndt})$
Carrying capacity (K) = population when $\frac{dN}{Ndt} = 0$  
K = $5.282*10^7 individuals$ 

\pagebreak
  
  
####c. Modified Plots 

```{r plot_modify, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(pop, aes(x = pop_mil, y = dNNdt)) +
  geom_point(aes(color=year, size=calories_pd), alpha = 0.8) +
  facet_wrap(~country, scales = "free") +
  labs(x = "Population Size (Millions of People)",
       y = "Per Capita Rate of Change",
       size = "Food Calories (Per Person Per Day)") +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom",
        legend.box = "vertical") + 
  scale_color_distiller(palette = "GnBu", direction = 1, 
                        name = "Year") 

```
  
`r fig_nums("all_color")`  

###Question 3 What Influences Growth    
  
     
```{r, include=FALSE}
# Packages for question 3

library(tidyverse)
library(stargazer)
library(tinytex)

```

#### Multiple Linear Regression  

```{r combined_regression, include=FALSE}

combined_model <- lm(dNNdt ~ years_ed + calories_pd, data=population)

# Years is significant p < 0.001 and calories is not  p = 0.549
```


```{r single_regression, include=FALSE}
########## Single Country Regressions 

#Brazil:  

brazil_model <- population %>% 
  filter(country == "Brazil") %>% 
  lm(dNNdt ~ years_ed + calories_pd, data = .)


# India:
india_model <- population %>% 
  filter(country == "India") %>% 
  lm(dNNdt ~ years_ed + calories_pd, data = .)


# Japan:
japan_model <- population %>% 
  filter(country == "Japan") %>% 
  lm(dNNdt ~ years_ed + calories_pd, data = .)

# Mexico: 
mexico_model <- population %>% 
  filter(country == "Mexico") %>% 
  lm(dNNdt ~ years_ed + calories_pd, data = .)

# South Korea:
skorea_model <- population %>% 
  filter(country == "South Korea") %>% 
  lm(dNNdt ~ years_ed + calories_pd, data = .)
```



```{r stargazer_country, include = FALSE}

table <- stargazer(combined_model, brazil_model, india_model, japan_model, mexico_model, skorea_model,
                      title = "Regression Results.",
                      type = "latex",
                      dep.var.labels = c("Per Capita Rate of Change (dN/Ndt)"),
                      column.labels = c("All", "Brazil", "India", "Japan", "Mexico", "South Korea"),
                      covariate.labels = c("Years of Education", 
                                           "Food Calories Per Person Per Day"),
                      digits=7,
                      header = FALSE)

```




```{r stargazer_results, results = 'asis', echo = FALSE, message = FALSE}
# formatting for final table 


table <- gsub("\\begin{tabular}","\\resizebox{1.0\\textwidth}{!}{\\begin{tabular}", table,fixed=T)
table <- gsub("\\end{tabular}","\\end{tabular}}", table,fixed=T)
cat(table)

```

\pagebreak


*Description of Variables*  
</p>

* Years of Education:
    + All:  For every increase in one year of education, we would expect the rat of change for the population to decrease by 0.0022489 if all other variables in the model remain the same.   
    + Brazil: For every increase in one year of education, we would expect the rate of change of the for the population to decrease by 0.0024205 if all other variables in the model remain the same.  
    + India: For every increase in one year of education, we would expect the rate of change of the for the population to decrease by 0.0015234 if all other variables in the model remain the same.   
    + Japan:For every increase in one year of education, we would expect the rate of change of the for the population to decrease by 0.0038341 if all other variables in the model remain the same.   
    + Mexico: For every increase in one year of education, we would expect the rate of change of the for the population to decrease by 0.0022801 if all other variables in the model remain the same.  
    + South Korea: For every increase in one year of education, we would expect the rate of change of the for the population to decrease by 0.0027568 if all other variables in the model remain the same.   
    
</p>  
  
* Food Calories: 
    + All: For every additional calorie consumed per person per day, we would expect the rate of change of the population to decrease by 0.0000001, all other factors remaining equal.   
    + Brazil: For every additional calorie consumed per person per day, we would expect the rate of change of the population to decrease by 0.0000012, all other factors remaining equal.  
    + India: For every additional calorie consumed per person per day, we would expect the rate of change of the population to decrease by 0.0000004, all other factors remaining equal.   
    + Japan: For every additional calorie consumed per person per day, we would expect the rate of change of the population to decrease by 0.0000020, all other factors remaining equal.  
    + Mexico: For every additional calorie consumed per person per day, we would expect the rate of change of the population to decrease by 0.0000053, all other factors remaining equal.  
    + South Korea: For every additional calorie consumed per person per day, we would expect the rate of change of the population to decrease by 0.0000004, all other factors remaining equal.  

