---
title: "ESM 201 HW 1"
author: "Sara Orofino"
date: "1/30/2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(kableExtra)
library(knitr)
library(stargazer)
library(tinytex)
library(captioner)

population <- read_csv("assig1_data.csv")
```

```{r fig_caps}
# set up for figure captions:
fig_nums <- captioner()
fig_nums("mex_percap", "Mexico's Actual vs. Calculated Per Capita Rate of Change (dN/Ndt). The graph depicts the actual vs. calculated per capita rate of change with a one-to-one trendline for Mexico from 1963 to 2003.")
fig_nums("all_percap", "Per Capita for All")
```


###Question 1 Calculate $\frac{dN}{Ndt}$ 

####a. Describe how annual population  are used to calculate $\frac{dN}{Ndt}$   

$\frac{dN}{Ndt}$ is the per capita rate of change of the population. Annual population data can be used to calculate the per capita rate of change (r) by taking the natural log of the number of individuals in the population at time t and divide by the number of individuals in the population at time t+1, multiplying it by $\frac{1}{(t+1)-t}$. In this case, the per capita rate of change would be the natural log of the number of individuals in the population in 1964 divided by the number of individuals in the population in 1963. For the difference from 1964 to 1963, the final term $\frac{1}{(t+1)-t}$ would be 1 and is not necessary to include in the equation. 

####b. Equations for calculating $\frac{dN}{Ndt}$  

* Brazil - $r = ln\left(\frac{81972001}{79602001}\right)$  

* India -  $r = ln\left(\frac{486639001}{476632001}\right)$   

* Japan -  $r = ln\left(\frac{96959001}{95929001}\right)$ 

* Mexico - $r = ln\left(\frac{43052001}{41715001}\right)$  

* South Korea - $r = ln\left(\frac{27767001}{27138001}\right)$    

```{r percap_ratechange, include=FALSE}

rates <- population %>% 
  filter(year %in% c(1963,1964)) %>% 
  select(country, year, population) %>% 
  spread(year, population) %>% 
  magrittr::set_colnames(value = c("country", "n0", "nt")) %>% 
  mutate(dNNdt_calc = log(nt/n0),
         dNNdt = c(0.029338507, 0.020777870, 0.010679874, 0.031547910, 0.022913305))


```

  
    
```{r table1, echo=FALSE}

rate_table  <- rates %>%
  select(country, dNNdt_calc, dNNdt)

kable(rate_table, digits = 4, col.names = c("Country",  "Calculated dNNdt", "Actual dNNdt"),
      caption = "Table 1. Calculated vs. Actual Per Capita Rate of Change by Country") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "bordered")) 

```
  
  
   
####c. Calcuate $\frac{dN}{Ndt}$ for Mexico for all years 1963-2004

```{r mexico, echo=FALSE, message=FALSE, warning=FALSE}

mexico <- read_csv("mexico_data.csv") %>% 
  ggplot(aes(x = dNNdt_calc, y = dNNdt)) +
  geom_point() +
  geom_smooth(method="lm", se=FALSE, size = 0.5, color = "midnightblue") +
  labs(x = "Calculated Per Capita Rate of Change (dN/Ndt)", 
       y = "Actual Per Capita Rate of Change (dN/Ndt)") +
  scale_x_continuous(limits = c(0.010, 0.0325), breaks = seq(0.010,0.030, by=0.005),
                     expand = c(0,0)) +
  scale_y_continuous(limits = c(0.010,0.0325),breaks =  seq(0.010,0.030, by=0.005),
                     expand  = c(0,0)) +
  theme_light()

mexico
```
`r fig_nums("mex_percap")`    


###Question 2 Graph $\frac{dN}{Ndt}$ vs. N~t~

####a. For Each Country Plot $\frac{dN}{Ndt}$ vs. N~t~

```{r pop_graphs, echo=FALSE}

ggplot(population, aes(x = population, y = dNNdt)) +
  geom_line() +
  facet_wrap(~country, scales = "free") +
  labs(x = "Population Size (N)",
       y = "Per Capita Rate of Change (dN/Ndt)") + 
  theme_light() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  
  
```


####b. Estimating Carrying Capacity 

```{r k, include=FALSE, message=FALSE, warning=FALSE}
# estimate carrying capacity for South Korea and Brazil 

# South Korea:

skorea_carry <- population %>% 
  filter(country == "South Korea") %>% 
  lm(population ~ dNNdt, data = .)

summary(skorea_carry)
# Equation: Population = 5.282x10^7 - 1.054x10^9(dNNdt)
# dNNdt = 0 at k --> K = 5.282x10^7

# Brazil:
brazil_carry <- population %>% 
  filter(country == "Brazil") %>% 
  lm(population ~ dNNdt, data = .)

summary(brazil_carry)
# Equation: Population = 2.607x10^8 - 6.392x10^9(dNNdt)
# dNNdt = 0 at k --> K = 2.607x10^8

```

  
**Carrying Capacities:**
  
- **Brazil:**   
$Population = 2.607*10^8 - 6.392*10^9(\frac{dN}{Ndt})$        
Carrying capacity (K) = population when $\frac{dN}{Ndt} = 0$   
$K = 2.607*10^8$  
  
- **South Korea:**   
$Population = 5.282*10^7 - 1.054*10^9(\frac{dN}{Ndt})$  
Carrying capacity (K) = population when $\frac{dN}{Ndt} = 0$   
K = $5.282*10^7$ 
  
####c. Modified Plots 

```{r plot_modify, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(population, aes(x = population, y = dNNdt)) +
  geom_point(aes(color=year, size=calories_pd), alpha = 0.8) +
  facet_wrap(~country, scales = "free") +
  labs(x = "Population Size (N)",
       y = "Per Capita Rate of Change (dN/Ndt)",
       size = "Food Calories (Person/Day)") +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom",
        legend.box = "vertical") + 
  scale_color_distiller(palette = "GnBu", direction = 1, 
                        name = "Year") 

```

###Question 3 What Influences Growth  
  

####a. Combined Multiple Linear Regression   
     

```{r combined_regression, include=FALSE}

combined_model <- lm(dNNdt ~ country + years_ed + calories_pd, data=population)

# Reference country Brazil; years and calories both significant p < 0.001 
```

```{r stargazer_combined, results='asis', echo=FALSE}

combined_table <- stargazer(combined_model, 
                            title = "Table 2. Combined Regression Results",
                            align = TRUE,
                            dep.var.labels = c("Per Capita Rate of Change (dN/Ndt)"),
                            covariate.labels = c("India", "Japan", "Mexico", "South Korea",
                                                 "Years of Education", 
                                                 "Food Calories Per Person Per Day"),
                            no.space = TRUE,
                            digits=2,
                            type = "html",
                            header=FALSE)

```
  
*Description of Variables:*

* 

  
####b. Multiple Linear Regression by Country   
  

```{r single_regression, include=FALSE}
########## Single Country Regressions 

#Brazil:  

brazil_model <- population %>% 
  filter(country == "Brazil") %>% 
  lm(dNNdt ~ years_ed + calories_pd, data = .)


# India:
india_model <- population %>% 
  filter(country == "India") %>% 
  lm(dNNdt ~ years_ed + calories_pd, data = .)


# Japan:
japan_model <- population %>% 
  filter(country == "Japan") %>% 
  lm(dNNdt ~ years_ed + calories_pd, data = .)

# Mexico: 
mexico_model <- population %>% 
  filter(country == "Mexico") %>% 
  lm(dNNdt ~ years_ed + calories_pd, data = .)

# South Korea:
skorea_model <- population %>% 
  filter(country == "South Korea") %>% 
  lm(dNNdt ~ years_ed + calories_pd, data = .)
```

```{r stargazer_countries, echo=FALSE, results= 'asis'}

lm_table <- stargazer(brazil_model, india_model, japan_model, mexico_model, skorea_model,
                      title = "Table 3. Regression Results by Country",
                      align = TRUE,
                      dep.var.labels = c("Per Capita Rate of Change (dN/Ndt)"),
                      column.labels = c("Brazil", "India", "Japan", "Mexico", "South Korea"),
                      covariate.labels = c("Years of Education", 
                                           "Food Calories Per Person Per Day"),
                      no.space = TRUE,
                      digits=2,
                      type = "html",
                      header = FALSE)
```




